My title

### 2024-07-08 paper/paper.qmd
updating heatmap

By isolating all three party coalitions, the it is easier to notice trends in the seat projections in the months leading up to the election. 
Particularly for the NDA, the projections become more skewed and the confirmation bias

@@

@@

# Print summary of the linear model
summary(lm_model)

@@

# Print the summary table
print(summary_table)

# Check if the relationship is statistically significant (p-value < 0.05)
p_value <- summary(lm_model)$coefficients["`Sample size`", "Pr(>|t|)"]
if (!is.na(p_value) && p_value < 0.05) {
  cat("The relationship between Sample size and NDA seats is statistically significant (p < 0.05).\n")
} else {
  cat("The relationship between Sample size and NDA seats is not statistically significant (p >= 0.05).\n")
}

@@

>>>>

By isolating all three party coalitions, it is easier to notice trends in the seat projections in the months leading up to the election. Particularly for the NDA, the predictions became skewed and deviated further from a more realistic end result. The target starts as somewhat realistic in April 2023, given that the NDA won 353 seats in the 2019 General Elections, and the general opinion of the public seems to be that they will surpass this previous record with the next election. However, as the election date approached, the projections increased dramatically, peaking at over 400 seats. This suggests an overestimation of NDA's popularity or a possible bias in the polling data.
Conversely, INDIA being a relatively young coalition, sees dynamic changes in predictions and a general underconfidence as the underdog in the election. Initially, the projections are relatively stable but show a significant drop around October 2023, which indicates possible public skepticism or fluctuation in support. The coalition's seat projections are consistently below the actual outcome, highlighting the tendency of the polls to underestimate their potential.
The predictions for the Others seats (which notably included the Bahujan Samaj Party and the AIADMK) were the most accurate, mostly because these are highly regional parties, making it easier to predict the seats they would bag with confidence. The projections for Others show a steady decline, aligning closely with the actual results. This accuracy can be attributed to the strong regional bases and consistent voter support for these parties, which makes their outcomes more predictable compared to the broader national coalitions.

@@

It is generally understood that when the sample size in a study increases, it represents a larger subset of the population and should therefore be more accurate about the average since it reduces sample-to-sample variability. Even if polls are conducted by different agencies, it is reasonable to assume that larger sample sizes should yield more accurate seat projections.

To test this assumption, a linear regression model was created using rstanarm to see if there is a correlation between sample size and the accuracy of the opinion poll predictions. One particular data point stood out as an outlier due to its exceptionally large sample size: the TV9-Bharatvarsh poll conducted in April 2024, with a sample size of 2,500,000 voters. A graph excluding this point was created to better visualize the data points clustering between 0 and 500,000.

The results showed no significant correlation between sample size and the accuracy of seat projections. The p-value was 0.6, which is far greater than the standard threshold of 0.05 for statistical significance. Interestingly, the trendline suggested that smaller sample sizes were more accurate in their predictions, contradicting the initial assumption. In other words, as more participants were included to represent a subset of the population more accurately, the results skewed and varied further away from the true value. A few reasons for the same are discussed in section .

@@

# Print summary of the linear model
#summary(lm_model)

@@

# Print the summary table
#print(summary_table)

# Check if the relationship is statistically significant (p-value < 0.05)
#p_value <- summary(lm_model)$coefficients["`Sample size`", "Pr(>|t|)"]
#if (!is.na(p_value) && p_value < 0.05) {
#  cat("The relationship between Sample size and NDA seats is statistically significant (p < 0.05).\n")
#} else {
#  cat("The relationship between Sample size and NDA seats is not statistically significant (p >= 0.05).\n")
#}

@@

## The exit polls and godi media

As mentioned in previous sections, 'godi media' and inflated opinion poll numbers lead to uncharacteristically high predictions for a right-wing landslide victory in India. After scouring multiple media articles, social media opinions, and analyzing general prevalent public opinions through conversations (compiled in the resources folder in __), a pro-right rating can be collated for the media houses. Here is an excerpt from one such source:

Prominent among these pro-BJP/RSS news outlets are English television channels Republic TV, India Today, Wion, and CNN-News18, and Hindi TV channels Zee News, India TV, AajTak, ABP News, Republic Bharat, Sudershan News, News Nation, and News24 (India).

Conversely, DB Live is widely understood to be what many refer to as a 'mouthpiece of the Indian National Congress'. It is one of the few channels in a rapidly declining list to show outward criticism of the ruling government's policies. Out of 180 countries, India ranked 161 in the Reporters Without Borders' World Press Freedom index in 2023, down 21 positions since Modi became prime minister in 2014. According to media watchdog NewsLaundry, an analysis of 429 primetime segments by 6 anchors in the run-up to the polls revealed that 52% of on-air time was spent criticizing the opposition, while only a small percentage of that time was spent on topics like jobs, education, and social issues.

Heatmap Analysis
To visually represent the relationship between the pro-NDA bias and seat projections, a heatmap of the Pro-NDA score alongside the seat projections of various news agencies is presented below. This graph should be approached with caution due to minimal secondary research available to support it, but it provides a valuable lens to consider when examining media bias and its potential impact on opinion poll results.

```{r}
#| echo: false
#| warning: false
#| label: fig-oprrinionffpoll
#| fig-cap: "Opinion poll Seat Prosdfjections"

media_data <- data.frame(
  Media_Outlets = c("ABP News-CVoter", "Agni News Services", "Dainik Bhaskar", "DB Live", 
                    "India Today-Axis My India", "India News-Dynamics", "India TV-CNX", 
                    "NDTV-Jan Ki Baat", "News18-CNBC", "News 24-Today's Chanakya", 
                    "News Nation", "Republic TV-Matrize", "Republic TV-PMarq", 
                    "Times Now-ETG", "TV9 Bharatvarsh"),
  NDA_Predictions = c(368, 242, 316, 221, 381, 371, 386, 377, 362, 400, 360, 360, 359, 358, 346),
  Pro_BJP_Score = c(7, 2, 4, 1, 9, 6, 8, 7, 5, 10, 3, 9, 9, 8, 7)  
)

# Sort data by Pro_BJP_Score
media_data <- media_data[order(media_data$Pro_BJP_Score, decreasing = TRUE), ]

# Create the heat map
ggplot(media_data, aes(x = reorder(Media_Outlets, Pro_BJP_Score), y = "NDA Prediction")) +
  geom_tile(aes(fill = NDA_Predictions), color = "white") +
  scale_fill_gradient(low = "blue", high = "orange", name = "NDA Predictions") +
  labs(title = "Heat Map of NDA Seat Predictions by Media Outlet",
       x = "Media Outlets",
       y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
